---
title: "p8105_hw2_rz2570"
author: "Ruilian Zhang"
date: "10/2/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```



# Problem 1

```{r}
# read and clean Mr. Trash Wheel data
trash_wheel_df = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-starts_with("x")) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))
```

```{r}
# read and clean precipitation data for 2018 and 2019
precipitation_2018 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  rename(total_precipitation = total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precipitation_2019 = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  rename(total_precipitation = total) %>% 
  mutate(year = 2019) %>% 
  relocate(year)

# combine precipitation datasets and convert month to a character variable
precipitation_df = bind_rows(precipitation_2018, precipitation_2019) %>% 
  mutate(month = month.name[month])
```

* In trash_wheel_df, there are `r nrow(trash_wheel_df)` observations. In precipitation_df, there are `r nrow(precipitation_df)` observations. 
* For example, in trash_wheel_df, the rounded number of sports balls in dumpster 1 is 7 and the number of cigarette butts is 126000 in May 2014. In precipitation_df, the total precipitation in June 2018 is 4.77 inch, and in June 2019 is 0.42 inch.
* For available data, the total precipitation in 2018 is `r sum(precipitation_2018$total_precipitation)`. The median number of sports balls in a dumpster in 2019 is `r median(trash_wheel_df$sports_balls)`.



# Problem 2

```{r}
# clean the data in pols-month.csv
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.numeric(year)) %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = ifelse(prez_gop == "1", "gop", "dem")) %>% 
  select(-c(prez_dem, prez_gop)) %>% 
  select(-day)
```

```{r}
# clean the data in snp.csv
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year"),sep = "/") %>% 
  select(-day) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = year + 2000) %>% 
  mutate(month = as.numeric(month)) %>%
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)
```

```{r}
# clean the data in unemployment.csv
unemployment_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "percentage"
  )
```

```{r}
# join the datasets by merging snp into pols
pols_merge = left_join(pols_df, snp_df)

# merge unemployment into the result and rename the variables
pols_merge = mutate(pols_merge, month = str_to_lower(month)) %>% 
  mutate(month = substr(month, 1, 3))

pols_result = left_join(pols_merge, unemployment_df) %>% 
  rename(
    president_party = president,
    close_stock_index = close,
    unemployment_percentage = percentage
  )
```

```{r, eval = FALSE}
# have a look at the resulting dataset
skimr::skim(pols_result)
```

* Dataset **pols_df** contains 822 observations of 9 variables related to the number of national politicians who are democratic or republican at any given time. The **snp_df** contains 787 observations of 3 variables, which are the date of observation and the closing values of the S&P stock index on the associated date. The **unemployment_df** contains 816 observations of 3 variables, which are the date of observation and the percentage of unemployment on the associated date. 
* The resulting **dataset pol_result** has 822 rows and 11 columns, containing all the variables from above datasets. The range of year lists from 1947 to 2015. Some key variables are  *president_party* (whether the president was democratic or republican), *close_stock_index* (the closing values of the S&P stock index) and *unemployment_percentage* (the percentage of unemployment).



# Problem 3

```{r}
# import and clean Popular_Baby_Names.csv
names_df = read_csv("data/Popular_Baby_Names.csv", col_types = "dffcdd") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_upper(childs_first_name))

# look at the variable of naming inconsistency and recode
levels(names_df$ethnicity)

names_df = mutate(
  names_df,
  ethnicity = recode(ethnicity,
                     "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                     "BLACK NON HISP" = "BLACK NON HISPANIC",
                     "WHITE NON HISP" = "WHITE NON HISPANIC"))

# remove duplicated rows
names_df = distinct(names_df)
```

```{r}
# produce a table showing the rank in popularity of the name “Olivia” as a female baby over time

olivia_df = filter(
  names_df,
  gender == "FEMALE",
  childs_first_name == "OLIVIA"
  ) %>% 
  select(-c(childs_first_name, gender, count))
  
olivia_table = pivot_wider(
  olivia_df,
  names_from = "year_of_birth",
  values_from = "rank"
)
```

```{r}
# produce a similar table showing the most popular name among male children over time
male_name_df = filter(names_df, gender == "MALE", rank == "1")
fmale_name_df = filter(names_df, gender == "FEMALE", rank == "1")
```

