---
title: "p8105_hw2_rz2570"
author: "Ruilian Zhang"
date: "10/2/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```



# Problem 1

```{r}
# read and clean Mr. Trash Wheel data
trash_wheel_df = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-starts_with("x")) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))
```

```{r}
# read and clean precipitation data for 2018 and 2019
precipitation_2018 = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  rename(total_precipitation = total) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precipitation_2019 = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na() %>% 
  rename(total_precipitation = total) %>% 
  mutate(year = 2019) %>% 
  relocate(year)

# combine precipitation datasets and convert month to a character variable
precipitation_df = bind_rows(precipitation_2018, precipitation_2019) %>% 
  mutate(month = month.name[month])
```

In trash_wheel_df, there are `r nrow(trash_wheel_df)` observations. In precipitation_df, there are `r nrow(precipitation_df)` observations.  
For example, in trash_wheel_df, the rounded number of sports balls in dumpster 1 is 7 and the number of cigarette butts is 126000 in May 2014. In precipitation_df, the total precipitation in June 2018 is 4.77 inch, and in June 2019 is 0.42 inch.  
For available data, the total precipitation in 2018 is `r sum(precipitation_2018$total_precipitation)`.  
The median number of sports balls in a dumpster in 2019 is `r median(trash_wheel_df$sports_balls)`.



# Problem 2

```{r}
# clean the data in pols-month.csv
pols_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  mutate(president = ifelse(prez_gop == "1", "gop", "dem")) %>% 
  select(-c(prez_dem, prez_gop)) %>% 
  select(-day)
```

```{r}
# clean the data in snp.csv
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year"),sep = "/") %>% 
  select(-day) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = year + 2000) %>% 
  mutate(month = as.numeric(month)) %>%
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month)
```

```{r}
unemployment_df = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    cols = jan:dec,
    names_to = "month",
    values_to = "percentage"
  )
```

